---
title: "Week 1 notes"
author: "Sofia Borges"
format: html
---
https://umgcccfcsr.github.io/CytometryInR/course/01_InstallingRPackages/

### To see which packages are attached an installed, run SessionInfo
```{r}
sessionInfo()
```

### installation of packages only needs to be done once. But attaching from the library needs to be done every time we want to use a package.
```{r}
install.packages("dplyr")
```

### installation needs "", but library (to use the package) does not. For CRAN repository.
```{r}
library(dplyr)
sessionInfo()
```

### not sure it will work since my computer and R version is older than dplyr##
```{r}
?select()
```

### When I hover over select nothing shows up?? Only in the console. It's supposed to show the packages###
```{r}
?ggplot
```

### Every package needs to be attached/loaded with (library()) before working with it. they then don't need to be deatached to stop working.### 
### There is a workaround this, using 2 :, such as ?dplyr::select()###
### How to remove the package (if there are dependencies we don't want to use), such as
```{r}
remove.packages("dplyr")
```

### REPOSITORIES - CRAN, Bioconductor, GitHub. The packages need to be installed from the correct repository.
### To access bioconductor
```{r}
install.packages("BiocManager")
```

### To install things on BioConductor, use install() instead of install.packages which is from CRAN
```{r}
?install()
```
```{r}
library(BiocManager)
```
### Example of a package for flow cytometry data QC
```{r}
install("PeacoQC")
```
```{r}
library(PeacoQC)
```
### Check if it has been attached to the local environment
```{r}
sessionInfo()
```

## Check what packages are installed and not loaded yet
```{r}
installed.packages()
```
### For installing from GitHub, from someone's repository - many don't have updates and may be abandoned. The intermediate package for GitHub is called remotes.
```{r}
install.packages("remotes")
```
```{r}
library(remotes)
```
### to install packages it's also different from GitHub install_github
```{r}
?install_github
```
### installing packages from GitHub, use the GitHub username followed by a "/" and then the name of the package repository
```{r}
install_github("hally166/flowSpectrum")
```
### TROUBLESHOOTING

### error: dependency is not available # it usually explains what needs to be installed
### Checking Packages VERSIONS
### Check package versions with sessionInfo or like this
```{r}
packageVersion("PeacoQC")
```
### To cite a package in a paper, and see the package version: 
```{r}
citation("PeacoQC")
```
### Installation of specific versions of packages, specifying the package, version and link of the repository, from CRAN
```{r}
remotes::install_version("ggplot2", version = "3.5.2", repos = "https://cloud.r-project.org")
```
### Versions of packages from GitHub usually don't follow the same formality in software versions. Or if the developer does not implement releases, the git commit can be provided for example
```{r}
remotes::install_github("DavidRach/Luciernaga", ref = "v0.99.7")
remotes::install_github("DavidRach/Luciernaga", ref = "8d1d694")
```
### Documentation and Websites for finding info on packages if they don't have or have few information with the HELP (?) info.### find on website and for github, look for the specific github repository page.###
Homework
```{r}
install.packages("tidyverse")
```
```{r}
install.packages("tidyverse")
```
```{r}
BiocManager::install("flowCore")
```
```{r}
BiocManager::install("flowWorkspace")
```
```{r}
install.packages("stringr")
```
```{r}
install("ggcyto")
```
```{r}
install.packages("tidyr")
```
```{r}
install("openCyto")
```
```{r}
install("FlowSOM")
```
```{r}
install.packages("xml2")
```
```{r}
install.packages("devtools")
```
```{r}
library(devtools)
install_github('saeyslab/CytoNorm')
```
```{r}
install.packages("lubridate")
```
```{r}
install.packages("httr2")
```
```{r}
BiocManager::install("flowGate")
```
```{r}
install.packages("plotly")
```
```{r}
BiocManager::install(c("flowCore", "Biobase", "sva"))
remotes::install_github("biosurf/cyCombine")
```
```{r}
BiocManager::install("CytoML")
```
```{r}
install.packages("Rtsne")
```
```{r}
install.packages("uwot")
```

### Take home problems (3) install other packages using the pak package